Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2019-05-08 paper/manuscript.tex
More edits on the relative free energy section

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Alchemical Free Energy Calculations: v\versionnumber}
\author[7]{Bryce K. Allen}
\author[1*]{John D. Chodera}
\author[2*]{Antonia S. J. S. Mey}
\author[2*]{Jenke Scheen}
\author[2*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[1]{Levi N. Naden}
\author[4*]{Samarjeet Prasad}
\author[5*]{Julia E. Rice}
\author[1,9]{Andrea Rizzi}
\author[6*]{Michael Shirts}
\author[10*]{Huafeng Xu}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Different steps required in planning and running a free energy calculation   %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What can be expected from alchemical simulations? -- Step 0}
\label{sec:step0}
\todo[inline, color={green!20}]{@everyone please review this section}
Step 0 is essentially about initially counting the cost of your
project: Can you even hope to tackle the problem you are thinking
about attempting given available resources and, if successful, will it
be worth it?
\subsection*{How accurate are current alchemical free energy calculations?}
Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. However, the domain of applicability is a significant concern~\cite{Sherborne2016, Cournia2017}, especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

@@

Here we include a simple estimate of the value provided by alchemical
free energy calculations in lead optimization.  Let $P(\Delta\Delta
G)$ be the probability distribution of the changes in the binding free
energies of a new set of molecules during one round of lead
optimization, and let $P(\Delta\Delta G^\dagger|\Delta\Delta G)$ be the
conditional probability of the binding free energy change computed by
the free energy calculations, $\Delta\Delta G^\dagger$, given the actual
change $\Delta\Delta G$.  The latter conditional probability can be modeled
by a normal distribution
\begin{equation}
  P(\Delta\Delta G^\dagger|\Delta\Delta G) = \frac{1}{\sqrt{2\pi\sigma^2}}
  \exp\left(-\frac{(\Delta\Delta G^\dagger - \Delta\Delta G)^2}{2\sigma^2}\right)
  \label{eqn:free-energy-distribution}
\end{equation}
where $\sigma$ signifies the accuracy of free energy calculations.
Here we assume that there is no systematic bias in the free energy
calculations, i.e., on average, the free energy change computed by
free energy calculations agrees with the actual free energy change.

In lead optimization guided by free energy calculations, we only
synthesize and experimentally test molecules that are predicted to
have favorable free energy changes.  We are thus interested in how
often that a molecule predicted to bind stronger actually turns out to
bind stronger.  In other words, we are interested in the conditional
probability
\begin{equation}
  P(\Delta\Delta G<0|\Delta\Delta G^\dagger<0)
  \label{eqn:true-positive}
\end{equation}

For illustrative purposes, we assume that the actual changes in the
binding free energies for a set of new molecules also follow normal
distribution, that the standard deviation in the changes is $RT\ln 5$
(corresponding to a 5-fold change in the binding affinities), and that
1 in 10 new molecules have increased binding affinity ($\Delta\Delta G
\leq 0$).  Under such assumptions, the conditional probability in
Eq.~\ref{eqn:true-positive} can be easily computed.  If the accuracy
of free energy calculations is $\sigma = 1$kcal/mol, $P(\Delta\Delta
G<0|\Delta\Delta G^\dagger<0) = 0.35$, which means that out of every
10 molecules selected for predicted favorable free energy change, on
average 3.5 molecules will have actual favorable free energy change.
In other words, selection by free energy calculations yields 3.5 times
more molecules of improved affinities than selection without free
energy calculations.
\todo[inline, color={green!20}]{@DM: A plot of this might be useful. I believe you have used/generated one of these before.}

Available computational resources and timescales of motion also factor
into this initial analysis. An individual free energy calculation
involves simulations at many different intermediate states (perhaps
20-40 or more) and each of these must typically be long enough to
capture the relevant motions in the system. If such motions are
microsecond events or longer, the computational cost of running 20-40
microsecond or longer simulations for each of $N$ ligands may become
prohibitive, or at least should be carefully considered. And, are
available computational resources sufficient that throughput will be
reasonable compared to needs of experimental collaborators working on
this system? How many ligands ($N$) can you afford to handle given
your computational resources?

@@

This analysis should be done up front, part of ``counting the cost''
of involvement in a particular project. In some cases, the result of
this analysis may be a conclusion that free energy calculations will
certainly not be feasible for the proposed problem.
\todo[inline, color={green!20}]{@everyone: Might be nice to do an example calculation X perturbation on GTX 1080s on a AWS cluster cost \$\$X. Synthesis costs \$\$Y. }

\subsection*{Is an explorative study what I want?}
An additional consideration is how much is known about your particular
target, ligand binding modes in the target, and any relevant motions
-- essentially, has it been studied enough to know whether it might be
suitable for free energy calculations? If the conclusion it is hardly
been studied, this is important to know going in as well as, if
initial calculations perform poorly, the effort may turn into an
attempt to understand the relevant sampling, force field, or system
preparation problems.

If you are unsure whether your project is feasible, one option may be
to conduct a short exploratory study to assess whether calculations
seem feasible at reasonable computational cost for just a very small
number of ligands. Sometimes this can be sufficient to get an initial
idea of feasibility and how accurate the calculations might be for the
proposed target.

\section{Simulation prerequisites -- Step 1}
\label{sec:step1}
\todo[inline,color={green!20}]{@everyone: please review}
The alchemical free energy protocols, which are discussed in section~\ref{sec:step2}, which can be used are tied in with what type of free energy you might want to compute, i.e. a free energy of binding or a free energy of hydration. Depending on the type of simulations different considerations have to be made for ligands, solvent, and host molecules (in the case of the estimation of free energies of binding).
\subsection*{Free energies of binding}
In principle in the limit of sufficient conformational sampling, the free energy changes estimated from an alchemical free energy calculation should be independent of the system's initial coordinates. However, in practice because simulations of thermodynamic states are of finite duration (typically currently 1-100 ns per state) this is only true for certain classes of alchemical free energy calculations such as relative or absolute free energies of hydration of small and relatively rigid organic molecules. Host-guest or protein-ligand complexes typically exhibit slowly relaxing degrees of freedom that exceed significantly the duration of an alchemical free energy calculation. It is therefore generally important to carefully select input coordinates to obtain satisfactory results. 
You might want to ask yourself the following questions before diving into your simulation setup. 
\begin{itemize}
    \item Do I have one or multiple good host structures? (e.g. a good resolution X-ray crystal of the protein target)
    \item Should I include burried waters, or other small molecules that can be found in an X-ray structure.
    \item Are my ligands part of a congeneric series? (simple R group substitutions around the same scaffold)
    \item Do I have information on one or all of the ligands binding sites (e.g. a X-ray structure)
\end{itemize}

@@

Generally binding modes within congeneric series are conserved (REFS?), however, exception exists(REF), and weaker binding compounds are more likely to adopt diverse binding modes (REFS). There may also be ambiguity about the relative orientation of substituents, a typical example are ortho- and meta-substituted 6-membered aromatic rings. While both positions are formally equivalent, a 180 degree flip of the ring may not occur sufficiently frequently during simulations. Another scenario may be equatorial and axial subsituted saturated rings (e.g. cyclohexane derivatives). This situation may be addressed by explicitly modelling different binding modes of the same ligand and combining later computed free energy differences for different binding modes into a relative free energies of binding.

@@

\subsection*{Free energies of hydration or partition coefficients}
Preliminary considerations necessary for using free energy methods to compute partition coefficients are much generally much more straight forward. Using even something such as bable to generate a 3D minimised structure of the solute you would like to compute a free energy of hydration or partition coefficient can be sufficient for preparation of these types of simulations. However, in these cases a careful choice of forcefields, as well as water models or organic solvents is essential. See for example XX.et al~\cite{xxx} for a good discussion on these choices.

\section*{What simulation protocol should I choose? -- Step 2}
\label{sec:step2}
\todo[inline, color=green!20]{Just a note as a reminder: Chirality!}
\subsection{Absolute and relative free energy calculations have some differences}

Alchemical free energy calculations can be grouped into two main categories, ``absolute'' and ``relative'' \footnote{The distinction is a bit of a misnomer, since both compute ratios of partition functions relative to another state, and neither computes an absolute free energy.}, which differ in whether they compute properties for a single molecule (absolute) or compare properties of different, usually closely related, molecules (relative).
To use binding as a concrete example, in absolute binding free energy calculations, one computes the binding free energy of a ligand to an individual receptor relative to a standard reference concentration.
In contrast, in relative binding free energy calculations, one compares the binding free energy of two related inhibitors to determine the potency difference.

Many protocol issues for alchemical calculations are common, but some are different between absolute and relative calculations, so before treating the common elements we treat the protocol differences.
\todo[inline]{JM: single/dual hybrid toplogies what are they?}

\subsubsection{Relative free energy calculations must select a topology and produce an atom mapping}

\paragraph{Topologies and atom mappings.} A critical first step in relative calculations is to select an approach to these calculations, determining whether to use a dual topology, single topology, or hybrid topology approach to relative calculations.
%Need figure?  Perhaps adapt http://www.alchemistry.org/wiki/Constructing_a_Pathway_of_Intermediate_States
The distinction between these can be illustrated by considering a hypothetical transformation from molecule A to molecule B, where both atoms share a common substructure but differ in which functional groups are present; e.g. consider a transformation of ethane (A) to methanol (B).
In this case the common substructure is at least CH3, though perhaps may be larger depending on how it is defined, as we discuss below.
In single topology calculations, the overall transformation is set up to involve as few additional atoms as possible, so ethane would be typically changed into methanol by changing two of the protons into non-interacting atoms called ``dummy atoms'' (retaining their bonded interactions but not interacting with the rest of the system) and the connected carbon mutated into an oxygen (with an associated change in the C-H bond parameters as the atoms change to an O-H).
Thus in a single topology calculation, atoms may change their type so relatively few dummy atoms are created.
In contrast, in a dual topology free energy calculation, no atoms are allowed to change type [ref Shirts book chapter in Computational Drug Discovery and Design] so the ethane to methanol transformation involves starting with ethane plus two non-interacting dummy atoms, then passing through an intermediate state where atoms which are becoming dummy atoms or ceasing being dummy atoms are partially interacting (this state may or may not be well defined [ref Mobley perspective]), and culminating in a state where methane is present along with three additional dummy atoms which were previously a corresponding methyl group of ethane.
%referneced http://www.alchemistry.org/wiki/Constructing_a_Pathway_of_Intermediate_States
Hybrid topology calculations have not seen much use [ref] but essentially consist of two absolute free energy calculations in opposite directions at the same time (turning one molecule off while turning the other on), and are best considered in that light.
At present, the most widely used approaches, such as in Schrodinger's FEP+[ref] and in FESetup[ref] (for which calculations may be planned with Lead Optimization Mapper (LOMAP) [refs]) seem to use single topology approaches, though some codes only support dual topology.
To our knowledge efficiency differences have not been thoroughly explored, though conventional wisdom would suggest that fewer dummy atoms are better [ref LOMAP paper/Mobley perspective].

Once a particular approach to the topology is selected, a crucial next step is to identify the common atoms which will not be perturbed.
Rigorously, this process essentially comprises a maximal common substructure (MCSS) search of the molecules involved to identify the common substructure -- though the parameters of the MCSS search will differ depending on whether single or dual topology calculations are planned.
Specifically, with a single topology approach in mind, atom types are allowed to change, so a permissive MCSS search can be done, whereas with dual topology a more strict search is required.
Some tools automate this process; for example, LOMAP can take a set of ligands and generate proposed pairings of molecules which are scored by their MCSS similarity and other properties [refs].
Schr\"{o}dinger's FEP+ planning tool is based on a version of LOMAP [ref].

@@

Single topology relative calculations, and calculations based on substructure searches, only work if in fact the ligands share a common substructure.
If no common substructure is shared, then essentially one ends up needing sophisticated dual or hybrid topology free energy calculations, where one would co-localize a pair of compounds in a binding site, exclude their interactions with one another, and compute the relative binding free energy by turning one molecule on from being dummy atoms while turning the other off.
To our knowledge no general pipeline for such calculations yet exists and this would likely remain a research problem.

@@

\paragraph{Constraints and relative free energy calculations.}
One issue which requires particular care is the use of constraints.
Commonly, bonds involving hydrogen are constrained to a fixed length to allow the use of longer timesteps.
However, in single topology relative free energy calculations, the atoms involved might be mutated to other atom types -- for example, in a mutation of methane to methanol, one hydrogen might become an oxygen atom.
Typical molecular dynamics engines are not set up to recognize this change, or at least not to correctly include contributions to the free energy from changing constraints/constraint length, so results for a transformation would usually be erroneous.
At present the most general solution to this problem is simply to avoid the use of constraints (and thus use a smaller timestep if necessary) in any relative free energy calculation involving a transformation of a constrained bond.

\subsubsection{Absolute free energy calculations must handle the standard state and use restraints}
\label{sec:standardstate-restraints}

@@

\subsection{Absolute and relative calculations must deal with some of the same issues}

\subsubsection{Structural definition of the bound state and weak binders}

@@

\subsubsection{The alchemical pathway is quite important \label{sec:important_path}}

@@

\subsubsection{THE BELOW STUFF SHOULD BECOME SUBSUBSECTIONS OR SUBSECTIONS OR...}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Alchemical Free Energy Calculations: v\versionnumber}
\author[7]{Bryce K. Allen}
\author[1*]{John D. Chodera}
\author[2*]{Antonia S. J. S. Mey}
\author[2*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[1]{Levi N. Naden}
\author[4*]{Samarjeet Prasad}
\author[5*]{Julia E. Rice}
\author[1,9]{Andrea Rizzi}
\author[2*]{Jenke Scheen}
\author[6*]{Michael Shirts}
\author[10*]{Huafeng Xu}

@@

%%%%%%%%%%%%
% Step 0   %
%%%%%%%%%%%%
\section{What can be expected from alchemical simulations? -- Step 0}
\label{sec:step0}
\todo[inline, color={green!20}]{@everyone please review this section}
Step 0 is essentially about initially counting the cost of your
project: Can you even hope to tackle the problem you are thinking
about attempting given available resources and, if successful, will it
be worth it?
\subsection*{How accurate are current alchemical free energy calculations?}
Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. However, the domain of applicability is a significant concern~\cite{Sherborne2016, Cournia2017}, especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

@@

Here we include a simple estimate of the value provided by alchemical
free energy calculations in lead optimization.  Let $P(\Delta\Delta
G)$ be the probability distribution of the changes in the binding free
energies of a new set of molecules during one round of lead
optimization, and let $P(\Delta\Delta G^\dagger|\Delta\Delta G)$ be the
conditional probability of the binding free energy change computed by
the free energy calculations, $\Delta\Delta G^\dagger$, given the actual
change $\Delta\Delta G$.  The latter conditional probability can be modeled
by a normal distribution
\begin{equation}
  P(\Delta\Delta G^\dagger|\Delta\Delta G) = \frac{1}{\sqrt{2\pi\sigma^2}}
  \exp\left(-\frac{(\Delta\Delta G^\dagger - \Delta\Delta G)^2}{2\sigma^2}\right)
  \label{eqn:free-energy-distribution}
\end{equation}
where $\sigma$ signifies the accuracy of free energy calculations.
Here we assume that there is no systematic bias in the free energy
calculations, i.e., on average, the free energy change computed by
free energy calculations agrees with the actual free energy change.
%
In lead optimization guided by free energy calculations, we only
synthesize and experimentally test molecules that are predicted to
have favorable free energy changes.  We are thus interested in how
often that a molecule predicted to bind stronger actually turns out to
bind stronger.  In other words, we are interested in the conditional
probability
\begin{equation}
  P(\Delta\Delta G<0|\Delta\Delta G^\dagger<0)
  \label{eqn:true-positive}
\end{equation}
%
For illustrative purposes, we assume that the actual changes in the
binding free energies for a set of new molecules also follow normal
distribution, that the standard deviation in the changes is $RT\ln 5$
(corresponding to a 5-fold change in the binding affinities), and that
1 in 10 new molecules have increased binding affinity ($\Delta\Delta G
\leq 0$).  Under such assumptions, the conditional probability in
Eq.~\ref{eqn:true-positive} can be easily computed.  If the accuracy
of free energy calculations is $\sigma = 1$kcal/mol, $P(\Delta\Delta
G<0|\Delta\Delta G^\dagger<0) = 0.35$, which means that out of every
10 molecules selected for predicted favorable free energy change, on
average 3.5 molecules will have actual favorable free energy change.
In other words, selection by free energy calculations yields 3.5 times
more molecules of improved affinities than selection without free
energy calculations.
\todo[inline, color={green!20}]{@DM: A plot of this might be useful. I believe you have used/generated one of these before.}
%  
Available computational resources and timescales of motion also factor
into this initial analysis. An individual free energy calculation
involves simulations at many different intermediate states (perhaps
20-40 or more) and each of these must typically be long enough to
capture the relevant motions in the system. If such motions are
microsecond events or longer, the computational cost of running 20-40
microsecond or longer simulations for each of $N$ ligands may become
prohibitive, or at least should be carefully considered. And, are
available computational resources sufficient that throughput will be
reasonable compared to needs of experimental collaborators working on
this system? How many ligands ($N$) can you afford to handle given
your computational resources?
%

@@

This analysis should be done up front, part of ``counting the cost''
of involvement in a particular project. In some cases, the result of
this analysis may be a conclusion that free energy calculations will
certainly not be feasible for the proposed problem.
\todo[inline, color={green!20}]{@everyone: Might be nice to do an example calculation X perturbation on GTX 1080s on a AWS cluster cost \$\$X. Synthesis costs \$\$Y. }
%
\subsection*{Is an explorative study what I want?}
An additional consideration is how much is known about your particular
target, ligand binding modes in the target, and any relevant motions
-- essentially, has it been studied enough to know whether it might be
suitable for free energy calculations? If the conclusion it is hardly
been studied, this is important to know going in as well as, if
initial calculations perform poorly, the effort may turn into an
attempt to understand the relevant sampling, force field, or system
preparation problems.
%
If you are unsure whether your project is feasible, one option may be
to conduct a short exploratory study to assess whether calculations
seem feasible at reasonable computational cost for just a very small
number of ligands. Sometimes this can be sufficient to get an initial
idea of feasibility and how accurate the calculations might be for the
proposed target.
%
%
%
%%%%%%%%%%%%
% Step 1   %
%%%%%%%%%%%%
%
\section{Simulation prerequisites -- Step 1}
\label{sec:step1}
\todo[inline,color={green!20}]{@everyone: please review}
The alchemical free energy protocols, which are discussed in section~\ref{sec:step2}, which can be used are tied in with what type of free energy you might want to compute, i.e. a free energy of binding or a free energy of hydration. Depending on the type of simulations different considerations have to be made for ligands, solvent, and host molecules (in the case of the estimation of free energies of binding).
\subsection*{Free energies of binding}
In principle in the limit of sufficient conformational sampling, the free energy changes estimated from an alchemical free energy calculation should be independent of the system's initial coordinates. However, in practice because simulations of thermodynamic states are of finite duration (typically currently 1-100 ns per state) this is only true for certain classes of alchemical free energy calculations such as relative or absolute free energies of hydration of small and relatively rigid organic molecules. Host-guest or protein-ligand complexes typically exhibit slowly relaxing degrees of freedom that exceed significantly the duration of an alchemical free energy calculation. It is therefore generally important to carefully select input coordinates to obtain satisfactory results. 
You might want to ask yourself the following questions before diving into your simulation setup. 
    \item Do I have one or multiple good host structures? (e.g. a good resolution X-ray crystal of the protein target)
    \item Should I include burried waters, or other small molecules that can be found in an X-ray structure.
    \item Are my ligands part of a congeneric series? (simple R group substitutions around the same scaffold)
    \item Do I have information on one or all of the ligands binding sites (e.g. a X-ray structure)

@@

Generally binding modes within congeneric series are conserved (REFS?), however, exception exists(REF), and weaker binding compounds are more likely to adopt diverse binding modes (REFS). There may also be ambiguity about the relative orientation of substituents, a typical example are ortho- and meta-substituted 6-membered aromatic rings. While both positions are formally equivalent, a 180 degree flip of the ring may not occur sufficiently frequently during simulations. Another scenario may be equatorial and axial substituted saturated rings (e.g. cyclohexane derivatives). This situation may be addressed by explicitly modelling different binding modes of the same ligand and combining later computed free energy differences for different binding modes into a relative free energies of binding.

@@

\subsection*{Free energies of hydration or partition coefficients}
Preliminary considerations necessary for using free energy methods to compute partition coefficients are much generally much more straight forward. Using even something such as bable to generate a 3D minimised structure of the solute you would like to compute a free energy of hydration or partition coefficient can be sufficient for preparation of these types of simulations. However, in these cases a careful choice of forcefields, as well as water models or organic solvents is essential. See for example XX.et al~\cite{xxx} for a good discussion on these choices. 
\todo[inline, color=green!20]{Just a note as a reminder: Chirality!}
%
%
%
%%%%%%%%%%%%
% Step 1   %
%%%%%%%%%%%%
%
\section{What simulation protocol should I choose? -- Step 2}
\label{sec:step2}
Alchemical free energy calculations can be grouped into two main categories, ``absolute'' and ``relative'' \footnote{The distinction is a bit of a misnomer, since both compute ratios of partition functions relative to another state, and neither computes an absolute free energy.}, which differ in whether they compute properties for a single molecule (absolute) or compare properties of different, usually closely related, molecules (relative).
To use binding as a concrete example, in absolute binding free energy calculations, one computes the binding free energy of a ligand to an individual receptor relative to a standard reference concentration.
In contrast, in relative binding free energy calculations, one compares the binding free energy of two related inhibitors to determine the potency difference.
\subsection*{Absolute and relative free energy calculations have some differences}
Many protocol issues for alchemical calculations are common, but some are different between absolute and relative calculations, so before treating the common elements we treat the protocol differences.
\todo[inline, color=greeen!20]{JM: single/dual hybrid toplogies what are they? ASJSM: These have been addressed below}

\subsubsection*{Choices unique to relative free energy calculations}

\paragraph{Topologies} A critical first step in relative calculations is to select an approach to these calculations, determining whether to use a \emph{dual topology}, \emph{single topology}, or \emph{hybrid topology} approach to relative calculations.
\begin{figure}
    \includegraphics[width=0.95\linewidth]{paper/figures/Fig2_dummy.pdf}
    \caption{a) single topology b) dual topology Figure adapted from \url{http://www.alchemistry.org/wiki/Constructing_a_Pathway_of_Intermediate_States}}
    \label{fig:topology}
\end{figure} 
The distinction between these can be illustrated by considering a hypothetical transformation from molecule A to molecule B, where both atoms share a common substructure but differ in which functional groups are present; e.g. consider a transformation of benzene (A) to benzyl alcohol (B)~\ref{fig:topology}.
In this case the common substructure is the benzene ring, though perhaps may be larger depending on how it is defined, as we discuss below.
In single topology calculations, the overall transformation is set up to involve as few additional atoms as possible, so benzene would be typically changed into benzyl alcohol by changing one of the hydrogens into a carbon. To this new carbon you also find non-interacting atoms called ``dummy atoms'' (retaining their bonded interactions but not interacting with the rest of the system), which will become the two hydrogens attached to the new carbon and the OH group. Bond parameters as well as partial charges are adjusted accordingly. 
Thus in a single topology calculation, atoms may change their type so relatively few dummy atoms are created. This is illustrated in fig.~\ref{fig:topology} (a). 
In contrast, in a dual topology free energy calculation, no atoms are allowed to change type~\cite{Shirts2012}. This means that the benzene to benzyl alcohol transformation involves starting with benzene plus the non-interacting dummy atoms making up the hydroxy methyl group, then passing through an intermediate state where atoms which are becoming dummy atoms or ceasing being dummy atoms are partially interacting (this state may or may not be well defined~\cite{Mobley:2014:J.Comput.AidedMol.Des.}  and culminating in a state where benzyl alcohol is present along with the additional dummy atoms which was previously a corresponding hydrogen of the benzene. Fig.~\ref{fig:topology} (b) depicts how such a dual topology works.

Hybrid topology calculations have seen much less use but essentially consist of two absolute free energy calculations in opposite directions at the same time (turning one molecule off while turning the other on), and are best considered in that light~\cite{ [ref] }.
At present, the most widely used approaches, such as in Schrodinger's FEP+[ref] and in FESetup[ref] (for which calculations may be planned with Lead Optimization Mapper (LOMAP) [refs]) seem to use single topology approaches, though some codes only support dual topology.
To our knowledge efficiency differences have not been thoroughly explored, though conventional wisdom would suggest that fewer dummy atoms are better [ref LOMAP paper/Mobley perspective].

\paragraph{Atom mapping}
Once a particular approach to the topology is selected, a crucial next step is to identify the common atoms which will not be perturbed.
Rigorously, this process essentially comprises a maximal common substructure (MCSS) search of the molecules involved to identify the common substructure -- though the parameters of the MCSS search will differ depending on whether single or dual topology calculations are planned.
Specifically, with a single topology approach in mind, atom types are allowed to change, so a permissive MCSS search can be done, whereas with dual topology a more strict search is required.
\begin{figure}
    \includegraphics[width=0.95\linewidth]{paper/figures/Fig2_dummy.pdf}
    \caption{a) restrictive MCSS match b) ring breaking}
    \label{fig:mcss}
\end{figure}

There are different tools that allow the generation of MCSS matches as well as single topology input. A large number of software tools are nowadays available for computing MCSS matches, some of the rely on RDKit for this computation, such as LOMAP or FESetup and partially BioSimSpace, others such as fkckombu are standalone tools~\cite{rdkit, lomap, fesetup, biosimspace, fkcombu}. Schr\"{o}dinger's FEP+ planning tool is based on a version of LOMAP, also allows MCSS matching and planning single topology calculations between molecules~\cite{FEP+}.

@@

Single topology relative calculations, and calculations based on substructure searches, only work if in fact the ligands share a common substructure, e.g. are part of a congeneric series.
If no common substructure is shared, then essentially one ends up needing sophisticated dual or hybrid topology free energy calculations, where one would co-localize a pair of compounds in a binding site, exclude their interactions with one another, and compute the relative binding free energy by turning one molecule on from being dummy atoms while turning the other off.
To our knowledge no general pipeline for such calculations yet exists and this would likely remain a research problem. Using an absolute style free energy calculation instead seems a more promising approach in such a case.

@@

\paragraph{Perturbation maps}
As part of LOMAP and the FEP+ planning tool. Based on the input ligand series a perturbation map can be planned. Recent heuristics have shown the more connected the perturbation network the better, however there is away to optimise network structure while minimising the number of perturbations that need to be computed minimising the resulting computational cost~\cite{Mobely2019}. Sometimes the introduction of intermediates that are not part of the original congeneric series are essential to avoid ring breaking, or dealing with perturbations that would otherwise result in large numbers of atoms to be grown or disappearing. None commercial tools have underlying good heuristic but may fail with complicated input, needing user validation in particular when dealing with chiral compounds.

\paragraph{Constraints and relative free energy calculations.}
One issue which requires particular care is the use of constraints.
Commonly, bonds involving hydrogen are constrained to a fixed length to allow the use of longer timesteps.
However, in single topology relative free energy calculations, the atoms involved might be mutated to other atom types -- for example, in a mutation of methane to methanol, one hydrogen might become an oxygen atom.
Typical molecular dynamics engines are not set up to recognize this change, or at least not to correctly include contributions to the free energy from changing constraints/constraint length, so results for a transformation would usually be erroneous.
At present the most general solution to this problem is simply to avoid the use of constraints (and thus use a smaller timestep if necessary, usually of around 1 fs) in any relative free energy calculation involving a transformation of a constrained bond.

\subsubsection*{Absolute free energy calculations must handle the standard state and use restraints}
\label{sec:standardstate-restraints}

@@

\subsection*{Absolute and relative calculations must deal with some of the same issues}

\subsubsection*{Structural definition of the bound state and weak binders}

@@

\subsubsection*{The alchemical pathway is quite important \label{sec:important_path}}

@@

\subsubsection*{How long should I run my simulation for and what information should be saved?}
